---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# inops

The package *inops* provides operators which expand to sets, ranges and regular
expressions the functionality provided by the `%in%` infix operator, while keeping
their behavior close to the one of equality, inequality and comparison operators
(see dedicated section).

Operators can be categorized as follow :

* **detection operators** return a logical output (as `%in%` does)
* **subsetting operators** return a subset of the input.
* **assignment operators** overwrite the values of the input matches by the set, range, or regular expression. 

Install with:

```
remotes::install_github("moodymudskipper/inops")
```

## Detection operators

Detection operators are the most important as subsetting and assignment operators
are simple wrappers around them, they are of the form `%<in or out><suffix>%`.

### Set operations with `%in{}%`

`%in{}%` is very similar with `%in%` at first sight.

```{r}
library(rangeops)
8:11 %in{}% 1:10
```

The difference is that it behaves consistently with `==`, which means that it
can be used on object of different types or classes and the output will be of the
same shape.

For instance we can use `%in{}%` on data frames :

```{r}
df1 <- data.frame(a = 1:3, b = 2:4, c=letters[1:3])
df1 == 2
df1 %in% 2
df1 %in{}% 2
df1 %in{}% 2:3
```

Missing values are not turned to `FALSE` :

```{r}
NA == 1
NA %in% 1
NA %in{}% 1
c(1,NA, 3) %in{}% 1:10
```

All "in" operators of this package come with a "out" counterpart.

```{r}
c(1,NA, 3) %out{}% 1:10
```

We won't mention the other "out" operators in the rest of the document, nor the
behavior of operators on non atomic data datypes, for brievety, but they work
just the same for other *inops* operator, with the exception of `%out%`, 
the negation of `%in%` provided for convenience.

### Range operations with `%in[]%`, `%in[)%`, `%in(]%` and `%in()`

These check if values are in a given range, the rhs can be a simple pair of
values or a bigger set, in which case the range of these values will be considered.

`[` and `]` describe closed boundaries while `(` and `)` describe open boundaries.

```{r}
8:12 %in[]% c(9, 10, 6)
8:12 %in[)% c(9, 10, 6)
```

### Regex operations with `%in~%`, `%in~p%` and `%in~f%`

These operators are wrappers around `grepl()`, `%in~%` uses the default parameters
while `%in~p%` uses `perl = TRUE` and `%in~f%` uses `fixed = TRUE`.

```{r}
c("foo","bar","baz") %in~% c("^f","z$")
```

## Subsetting operators

All detection operators have a subsetting operator counterpart, they are of the form
`%[<in or out><suffix>%`.

`%in%` and `%out%` were given subsetting counterparts `%[in%` and `%[out%` as
well.

```{r}
8:12 %[in[)% c(9, 10, 6)
8:12 %[out% c(5, 7, 9, 11)
c("foo","bar","baz") %[in~% c("^f","z$")
```

## Assignment operators

All detection operators have an assignment operator counterpart, it is in practice
typed just the same, but is followed by an assignment.

`%in%` and `%out%` were also given assignment counterparts `%in%<-` and `%out%<-`,
as were also `==`, `!=`, `<`, `>`, `<=` and `=>`.

```{r}
x <- 1:10
x %in[]% c(4:6) <- 0
x

x <- c("foo","bar","baz")
x %in~% c("^f","z$") <- "!!!"
x

x <- 1:10
x < 3 <- 0
x
```

Note that to give an assignment counterpart to `<` we had to overload the `<<-`
operator, which explains the message when attaching the package. This doesn't
affect the behavior of the `<<-` assignments.
