---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# inops

Package implementing additional infix operators for R.

Implemented operators provide 3 distinct operations: **detection**, **subsetting**, and **replacement**.\
And work with 3 different value types: **sets**, **intervals**, and **regular expressions**.

Install using the `remotes` package:

```
remotes::install_github("moodymudskipper/inops")
```

## Operators

Introduction to operator behaviour and design.

### Form ###

All operators have the same form composed of two distinct parts: `%<operation><type>%`.

* `[operation]` specifies the performed operation and can be one of `in`, `out`, `[in`, `[out`.
* `[type]` specifies the type of operation and can be one of `{}`, `[]`, `()`, `[)`, `(]`, `~`, `~p`, `~f`.

To understand what each combination does see the table below.

### Behaviour ###

The operators implemented here try to be consistent with the default comparison operators like `==` and `<`.
Therefore in some scenarios their behaviour differs from `%in%`.
For instance:

1) `%in{}%` can be used on on data frames.

```{r}
df1 <- data.frame(a = 1:3, b = 2:4, c=letters[1:3])

df1 == 2

df1 %in% 2

df1 %in{}% 2

df1 %in{}% 2:3
```

2) missing values are not considered as not matching.

```{r}
NA == 1

NA %in% 1

NA %in{}% 1

c(1, NA, 3) %in{}% 1:10
```

### Operations ###

Operators permit three distinct operations:

1. Detect: `x %in{}% set`
2. Subset: `x %[in{}% set`
3. Replace: `x %in{}% set <- value`

## Operator List ##

Below is a full list of all the implemented operators along with their usage examples.

### Detection Operators ###

```
|-----------|-----------------------------------------------------------|-----------------------------|
|  Form     |                    Description                            |            Call             |
|-----------|-----------------------------------------------------------|-----------------------------|
| %in{}%    | is element inside a set                                   | x %in{}% set                |
| %in[]%    | is element inside a closed interval                       | x %in[]% interval           |
| %in()%    | is element inside an open interval                        | x %in()% interval           |
| %in[)%    | is element inside an interval open on the right           | x %in[)% interval           |
| %in(]%    | is element inside an interval open on the left            | x %in(]% interval           |
| %in~%     | does element match a regular expression                   | x %in~% pattern             |
| %in~p%    | does element match a regular perl expression              | x %in~p% pattern            |
| %in~f%    | does element match a regular fixed expression             | x %in~f% pattern            |
| %out%     | is element outside a set (same as ! x %in% y)             | x %out% set                 |
| %out{}%   | is element outside a set                                  | x %out{}% set               |
| %out[]%   | is element outside a closed interval                      | x %out[]% interval          |
| %out()%   | is element outside an open interval                       | x %out()% interval          |
| %out[)%   | is element outside an interval open on the right          | x %out[)% interval          |
| %out(]%   | is element outside an interval open on the left           | x %out(]% interval          |
| %out~%    | does element not match a regular expression               | x %out~% pattern            |
| %out~p%   | does element not match a regular perl expression          | x %out~p% pattern           |
| %out~f%   | does element not match a regular fixed expression         | x %out~f% pattern           |
|-----------|-----------------------------------------------------------|-----------------------------|
```

### Subsetting Operators ###

```
|-----------|-----------------------------------------------------------|-----------------------------|
|  Form     |                    Description                            |            Call             |
|-----------|-----------------------------------------------------------|-----------------------------|
| %[in{}%   | select elements inside a set                              | x %[in{}% set               |
| %[in[]%   | select elements inside a closed interval                  | x %[in[]% interval          |
| %[in()%   | select elements inside an open interval                   | x %[in()% interval          |
| %[in[)%   | select elements inside an interval open on the right      | x %[in[)% interval          |
| %[in(]%   | select elements inside an interval open on the left       | x %[in(]% interval          |
| %[in~%    | select elements matching a regular expression             | x %[in~% pattern            |
| %[in~p%   | select elements matching a regular perl expression        | x %[in~p% pattern           |
| %[in~f%   | select elements matching a regular fixed expression       | x %[in~f% pattern           |
| %[out%    | select elements outside a set                             | x %[out%  set               |
| %[out{}%  | select elements outside a set                             | x %[out{}%  set             |
| %[out[]%  | select elements outside a closed interval                 | x %[out[]% interval         |
| %[out()%  | select elements outside an open interval                  | x %[out()% interval         |
| %[out[)%  | select elements outside an interval open on the right     | x %[out[)% interval         |
| %[out(]%  | select elements outside an interval open on the left      | x %[out(]% interval         |
| %[out~%   | select elements not matching a regular expression         | x %[out~% pattern           |
| %[out~p%  | select elements not matching a regular perl expression    | x %[out~p% pattern          |
| %[out~f%  | select elements not matching a regular fixed expression   | x %[out~f% pattern          |
|-----------|-----------------------------------------------------------|-----------------------------|
```

### Replacement Operators ###

```
|-----------|-----------------------------------------------------------|-----------------------------|
|  Form     |                    Description                            |            Call             |
|-----------|-----------------------------------------------------------|-----------------------------|
| %==<-%    | change elements equal to the provided value               | x == element <- value       |
| %!=<-%    | change elements not equal to the provided value           | x != element <- value       |
| %><-%     | change elements greater than the provided value           | x > number <- value         |
| %<<-%     | change elements lower than the provided value             | x < number  <- value        |
| %>=<-%    | change elements greater or equal to the provided value    | x >= number <- value        |
| %<=<-%    | change elements lower or equal to the provided value      | x <= number <- value        |
| %in{}<-%  | change elements inside a set                              | x %in{}% set <- value       |
| %in[]<-%  | change elements inside a closed interval                  | x %in[]% interval <- value  |
| %in()<-%  | change elements inside an open interval                   | x %in()% interval <- value  |
| %in[)<-%  | change elements inside an interval open on the right      | x %in[)% interval <- value  |
| %in(]<-%  | change elements inside an interval open on the left       | x %in(]% interval <- value  |
| %in~<-%   | change elements matching a regular expression             | x %in~% pattern <- value    |
| %in~p<-%  | change elements matching a regular perl expression        | x %in~p% pattern <- value   |
| %in~f<-%  | change elements matching a regular fixed expression       | x %in~f% pattern <- value   |
| %out<-%   | change elements outside a set                             | x %out% set <- value        |
| %out{}<-% | change elements outside a set                             | x %out{}% set <- value      |
| %out[]<-% | change elements outside a closed interval                 | x %out[]% interval <- value |
| %out()<-% | change elements outside an open interval                  | x %out()% interval <- value |
| %out[)<-% | change elements outside an interval open on the right     | x %out[)% interval <- value |
| %out(]<-% | change elements outside an interval open on the left      | x %out(]% interval <- value |
| %out~<-%  | change elements not matching a regular expression         | x %out~% pattern <- value   |
| %out~p<-% | change elements not matching a regular perl expression    | x %out~p% pattern <- value  |
| %out~f<-% | change elements not matching a regular fixed expression   | x %out~f% pattern <- value  |
|-----------|-----------------------------------------------------------|-----------------------------|
```

## Notes ##

To give an assignment counterpart to `<` we had to overload the `<<-` operator, which explains the message when attaching the package.
This doesn't affect the behavior of the `<<-` assignments.

