---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(width=110)
```

# inops

Package implementing additional infix operators for R.

Implemented operators provide 3 distinct operations: **detection**, **subsetting**, and **replacement**.\
And work with 3 different value types: **sets**, **intervals**, and **regular expressions**.

Install using the `remotes` package:

```
remotes::install_github("moodymudskipper/inops")
```

## Examples ##

Simple examples for illustration.

---

Selecting flight records from the `flights` dataset that:

1. Departed and Landed between noon and 5 p.m.
2. Were not traveling to "LEX", "PSP", nor "HDN"
3. Travelled distance was either very short (below 100) or very long (above 3000)
4. Had a tail number starting with "N1" or "N3"

```{r, message=FALSE}
library(nycflights13)
library(dplyr)
library(inops)

flights %>%
  filter(dep_time %in()%  c(1200, 1700)) %>%
  filter(arr_time %in()%  c(1200, 1700)) %>%
  filter(dest     %out%   c("LEX", "PSP", "HDN")) %>%
  filter(distance %out[]% c(100, 3000)) %>%
  filter(tailnum  %in~%   c("^N1", "^N3")) %>%
  select(origin, dest, tailnum, dep_time, arr_time, distance)
```

---

Cleaning up `planes` dataset to:

1. Standardize names of "AIRBUS", "CANADAIR" and "MCDONNELL" manufacturers.
2. Obtain plane counts for each of the 3 manufacturers mentioned above.

```{r, message=FALSE}
library(nycflights13)
library(inops)

table(planes$manufacturer %[in~% c("AIRBUS", "CANADAIR", "MCDONNELL"))

planes$manufacturer %in~% "AIRBUS"    <- "AIRBUS"
planes$manufacturer %in~% "CANADAIR"  <- "CANADAIR"
planes$manufacturer %in~% "MCDONNELL" <- "MCDONNELL"

table(planes$manufacturer %[in~% c("AIRBUS", "CANADAIR", "MCDONNELL"))

```

---

## Operators

Introduction to operator behaviour and design.

### Form ###

All operators have the same form composed of two distinct parts: `%<operation><type>%`.

* `[operation]` specifies the performed operation and can be one of `in`, `out`, `[in`, `[out`.
* `[type]` specifies the type of operation and can be one of `{}`, `[]`, `()`, `[)`, `(]`, `~`, `~p`, `~f`.

To understand what each combination does see the table below.

### Behaviour ###

The operators implemented here try to be consistent with the default comparison operators like `==` and `<`.\
Therefore in some scenarios their behaviour differs from `%in%`.\
For instance:

1) `%in{}%` can be used on on data frames.

```{r}
df1 <- data.frame(a = 1:3, b = 2:4, c=letters[1:3])

df1 == 2

df1 %in% 2

df1 %in{}% 2

df1 %in{}% 2:3
```

2) missing values are not considered as not matching.

```{r}
NA == 1

NA %in% 1

NA %in{}% 1

c(1, NA, 3) %in{}% 1:10
```

### Operations ###

Operators permit three distinct operations:

1. Detect: `x %in{}% set`
2. Subset: `x %[in{}% set`
3. Replace: `x %in{}% set <- value`

## Operator List ##

Below is a full list of all the implemented operators along with their usage examples.

### Detection Operators ###

|  Form       |                    Description                            |            Call               |
|-------------|-----------------------------------------------------------|-------------------------------|
| `%in{}%`    | is element inside a set                                   | `x %in{}% set`                |
| `%in[]%`    | is element inside a closed interval                       | `x %in[]% interval`           |
| `%in()%`    | is element inside an open interval                        | `x %in()% interval`           |
| `%in[)%`    | is element inside an interval open on the right           | `x %in[)% interval`           |
| `%in(]%`    | is element inside an interval open on the left            | `x %in(]% interval`           |
| `%in~%`     | does element match a regular expression                   | `x %in~% pattern`             |
| `%in~p%`    | does element match a regular perl expression              | `x %in~p% pattern`            |
| `%in~f%`    | does element match a regular fixed expression             | `x %in~f% pattern`            |
| `%out%`     | is element outside a set (same as ! x %in% y)             | `x %out% set`                 |
| `%out{}%`   | is element outside a set                                  | `x %out{}% set`               |
| `%out[]%`   | is element outside a closed interval                      | `x %out[]% interval`          |
| `%out()%`   | is element outside an open interval                       | `x %out()% interval`          |
| `%out[)%`   | is element outside an interval open on the right          | `x %out[)% interval`          |
| `%out(]%`   | is element outside an interval open on the left           | `x %out(]% interval`          |
| `%out~%`    | does element not match a regular expression               | `x %out~% pattern`            |
| `%out~p%`   | does element not match a regular perl expression          | `x %out~p% pattern`           |
| `%out~f%`   | does element not match a regular fixed expression         | `x %out~f% pattern`           |

### Subsetting Operators ###

|  Form       |                    Description                            |            Call               |
|-------------|-----------------------------------------------------------|-------------------------------|
| `%[in{}%`   | select elements inside a set                              | `x %[in{}% set`               |
| `%[in[]%`   | select elements inside a closed interval                  | `x %[in[]% interval`          |
| `%[in()%`   | select elements inside an open interval                   | `x %[in()% interval`          |
| `%[in[)%`   | select elements inside an interval open on the right      | `x %[in[)% interval`          |
| `%[in(]%`   | select elements inside an interval open on the left       | `x %[in(]% interval`          |
| `%[in~%`    | select elements matching a regular expression             | `x %[in~% pattern`            |
| `%[in~p%`   | select elements matching a regular perl expression        | `x %[in~p% pattern`           |
| `%[in~f%`   | select elements matching a regular fixed expression       | `x %[in~f% pattern`           |
| `%[out%`    | select elements outside a set                             | `x %[out%  set`               |
| `%[out{}%`  | select elements outside a set                             | `x %[out{}%  set`             |
| `%[out[]%`  | select elements outside a closed interval                 | `x %[out[]% interval`         |
| `%[out()%`  | select elements outside an open interval                  | `x %[out()% interval`         |
| `%[out[)%`  | select elements outside an interval open on the right     | `x %[out[)% interval`         |
| `%[out(]%`  | select elements outside an interval open on the left      | `x %[out(]% interval`         |
| `%[out~%`   | select elements not matching a regular expression         | `x %[out~% pattern`           |
| `%[out~p%`  | select elements not matching a regular perl expression    | `x %[out~p% pattern`          |
| `%[out~f%`  | select elements not matching a regular fixed expression   | `x %[out~f% pattern`          |

### Replacement Operators ###

|  Form       |                    Description                            |            Call               |
|-------------|-----------------------------------------------------------|-------------------------------|
| `%==<-%`    | change elements equal to the provided value               | `x == element <- value`       |
| `%!=<-%`    | change elements not equal to the provided value           | `x != element <- value`       |
| `%><-%`     | change elements greater than the provided value           | `x > number <- value`         |
| `%<<-%`     | change elements lower than the provided value             | `x < number  <- value`        |
| `%>=<-%`    | change elements greater or equal to the provided value    | `x >= number <- value`        |
| `%<=<-%`    | change elements lower or equal to the provided value      | `x <= number <- value`        |
| `%in{}<-%`  | change elements inside a set                              | `x %in{}% set <- value`       |
| `%in[]<-%`  | change elements inside a closed interval                  | `x %in[]% interval <- value`  |
| `%in()<-%`  | change elements inside an open interval                   | `x %in()% interval <- value`  |
| `%in[)<-%`  | change elements inside an interval open on the right      | `x %in[)% interval <- value`  |
| `%in(]<-%`  | change elements inside an interval open on the left       | `x %in(]% interval <- value`  |
| `%in~<-%`   | change elements matching a regular expression             | `x %in~% pattern <- value`    |
| `%in~p<-%`  | change elements matching a regular perl expression        | `x %in~p% pattern <- value`   |
| `%in~f<-%`  | change elements matching a regular fixed expression       | `x %in~f% pattern <- value`   |
| `%out<-%`   | change elements outside a set                             | `x %out% set <- value`        |
| `%out{}<-%` | change elements outside a set                             | `x %out{}% set <- value`      |
| `%out[]<-%` | change elements outside a closed interval                 | `x %out[]% interval <- value` |
| `%out()<-%` | change elements outside an open interval                  | `x %out()% interval <- value` |
| `%out[)<-%` | change elements outside an interval open on the right     | `x %out[)% interval <- value` |
| `%out(]<-%` | change elements outside an interval open on the left      | `x %out(]% interval <- value` |
| `%out~<-%`  | change elements not matching a regular expression         | `x %out~% pattern <- value`   |
| `%out~p<-%` | change elements not matching a regular perl expression    | `x %out~p% pattern <- value`  |
| `%out~f<-%` | change elements not matching a regular fixed expression   | `x %out~f% pattern <- value`  |

## Notes ##

To give an assignment counterpart to `<` we had to overload the `<<-` operator, which explains the message when attaching the package.
This doesn't affect the behavior of the `<<-` assignments.

